---
title: "Economic Vulnerability"
author: "VH"
date: "6/20/2019"
output: 
  github_document: default
  html_document: default
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
#Setting root directory
knitr::opts_knit$set(echo = TRUE,
                     root.dir = rprojroot::find_rstudio_root_file())

#To run the file you'll need to install all the packages here, in this order

#Packages
#install.packages("dplyr") # new  version
#install.packages(c("snakecase", "janitor")) #tidy names
#install.packages("pillar") #Dependency for patchwork
#devtools::install_github("thomasp85/patchwork") #For multi-panel plots
#install.packages(c("osmdata", "ggmap"))
#install.packages(c("sp", "sf"))
library(tidyverse) 
library(janitor)   #Clean names
library(patchwork) #Multi panel plots
library(osmdata)   #OSM
library(ggmap)     #Tidy gg maps
library(purrr)     #Map functions
library(sp)        #Spatial forms
library(sf)        #Spatial forms

#Controlling figure output in markdown
knitr::opts_chunk$set(
#  fig.height =   
  fig.width = 6,
#  fig.asp = .5,
  out.width = "90%",
#  out.height = 
  cache = FALSE
)

#Set Theme for ggplot2
theme_set(theme_bw() + theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom"))

#Set Scientific notation output for knitr
options(scipen = 999999)
```

1. Define the Functions  

Before you do anything, restart your R session (under Session at the top). Then run the setup chunk above to initilize the packages you need and other output definitions for the .RMD document (don't worry about these for now). Run this entire chunk to define the functions for use below. You can verify that this worked if they show up in your global envirnment (top right), under functions.  

```{r read_osm, warning = FALSE}
#Fairfax read OSM to output tibble (tidyr df)
osm_to_df <- function(key = "leisure", value = "playground", type = "point") {
#Read in playground data
read.df <- getbb("fairfax county", format_out = "polygon") %>%
  opq() %>%
  add_osm_feature(key = key, value = value) %>%
  osmdata_sf()
    
    if (type == "point" & !is.null(read.df$osm_points)) {
      
      lat.long.df <- do.call(rbind, st_geometry(read.df$osm_points)) %>% 
        as_tibble() %>% 
        setNames(c("longitude", "latitude")) %>%
        mutate(object_id = 1:nrow(.) %>% as.factor()) %>%
        dplyr::select(object_id, everything())
      
    } else if (type == "polygon" & !is.null(read.df$osm_polygons)) {
      
      poly.list <- do.call(rbind, st_geometry(read.df$osm_polygons)) %>%
        lapply(as.tibble)
      npolys    <- length(poly.list)
      
      lat.long.df <- tibble(
        object_id = 1:npolys %>% as.factor(),
        coord.df  = poly.list
        ) %>%
        unnest(coord.df) %>%
        rename(
        longitude = lon,
        latitude  = lat
        )
      
    } else if (type == "line" & !is.null(read.df$osm_lines)) {
      
      line.list <- do.call(list, st_geometry(read.df$osm_lines)) %>%
        lapply(., rbind) %>%
        lapply(., as.tibble)
      nlines    <- length(line.list)
      
      lat.long.df <- tibble(
        object_id = 1:nlines %>% as.factor(),
        coord.df  = line.list
        ) %>%
        unnest(coord.df) %>%
        rename(
        longitude = lon,
        latitude  = lat
        )
    } else {
      ifelse(is.null(read.df$osm_points), stop("key/value incorrect or object type is empty"), 
             ifelse(!(type %in% c("point", "polygon", "line")), stop("type is not point, polygon, or line"),
                    stop("object type is empty, try another")))
    }
  
  return(lat.long.df)
}

#Base Map Function
fairfax.gg <- function() {
fairfax.box <- getbb("fairfax county")
fairfax.boundary <- getbb("fairfax county", format_out = "polygon") %>%
  as.tibble() %>%
  rename(longitude = `V1`, latitude = `V2`)

#Grab the map info (many varieties)
fairfax.map <- get_map(location = fairfax.box, source="stamen", maptype="watercolor", crop = TRUE)

#ggmap and ggplot map and boundary
ff.map <- ggmap(fairfax.map) +
  geom_polygon(data = fairfax.boundary, aes(x = longitude, y = latitude), colour = "black", size = 1, alpha = 0.1) +
  labs(
    x = "Longitude",
    y = "Latitude"
  )
  return(ff.map)
}

#Point Visualization
osm_point_plot <- function(ff.map, obj, value) {
  if (!is.data.frame(obj)) stop("OSM object is not a data frame")
  if (!is.ggplot(ff.map))  stop("Baseline Fairfax map is not a ggplot")
  ff.map +
  geom_point(data = obj, aes(x = longitude, y = latitude),
             size = 0.1, colour = "red", alpha = 0.25) + 
  geom_density_2d(data = obj, aes(x = longitude, y = latitude),
                  colour = "purple", alpha = 0.5) +
  labs(title = sprintf("Fairfax County %s", value))
  }

#Polygon Plot
osm_poly_plot <- function(ff.map, obj, value) {
  if (!is.data.frame(obj)) stop("OSM object is not a data frame")
  if (!is.ggplot(ff.map))  stop("Baseline Fairfax map is not a ggplot")
  ff.map +
  geom_polygon(data = obj, aes(x = longitude, y = latitude, group = object_id), 
               colour = "red", fill = "maroon", alpha = 0.5, size = 0.2) +
  labs(title = sprintf("Fairfax County %s", value))
  }

#Line Plot
osm_line_plot <- function(ff.map, obj, value) {
  if (!is.data.frame(obj)) stop("OSM object is not a data frame")
  if (!is.ggplot(ff.map))  stop("Baseline Fairfax map is not a ggplot")
  ff.map +
  geom_path(data = obj, aes(x = longitude, y = latitude, group = object_id), 
               colour = "red", alpha = 0.5, size = 0.5) +
  labs(title = sprintf("Fairfax County %s", value))
  }

#File Path Generator (for OSM specifically)
osm_filepath_gen <- function(data.folder, file.type, key, value, type) {
  date <- Sys.Date() %>% as.character() %>% str_replace_all("-", "_")
  file.type <- "csv"
  data.folder <- "original/osm"
  file.path <- sprintf("./data/%s/%s_%s.%s", 
                     data.folder,
                     date, 
                     paste(key, value, type, sep = "_"),
                     file.type)
  return(file.path)
}

```


```{r}
ff.map <- fairfax.gg()
play.df <- osm_to_df(key = "leisure", value = "playground", type = "point")
playground.gg <- osm_point_plot(ff.map, play.df, "Playgrounds")  
playground.gg  
play.df$object_id %>% unique() %>% length()

```

Playground Observations: No way to cross check playgrounds on GMAPS!...
But, overall the points appear evenly spread & useable  
Use: Yes

```{r}

ff.map <- fairfax.gg()
play.df <- osm_to_df(key = "leisure", value = "park", type = "polygon")
playground.gg <- osm_poly_plot(ff.map, play.df, "Parks")  
playground.gg  
play.df$object_id %>% unique() %>% length()

```
OSM definition: "Open space for recreational use, usually designed and in semi-natural state with grassy areas, trees and bushes." 

Park Observation:Useful but accuracy questionable -> We may need to overlay the nature reserve/national park/garden maps to have a similar map? Cross check with Fairfax website data:

https://data-fairfaxcountygis.opendata.arcgis.com/datasets/county-parks?geometry=-78.124%2C38.645%2C-76.197%2C39.019

Use: Yes

```{r}

ff.map <- fairfax.gg()
play.df <- osm_to_df(key = "leisure", value = "swimming_pool", type = "point")
playground.gg <- osm_point_plot(ff.map, play.df, "Swimming Pools")  
playground.gg  
play.df$object_id %>% unique() %>% length()

```

Pool Observation: Looks decent. 

We can compare acuracy with Fairfax website data: https://data-fairfaxcountygis.opendata.arcgis.com/datasets/community-pools?geometry=-78.05%2C38.674%2C-76.124%2C39.048

Use: Yes

```{r}

ff.map <- fairfax.gg()
play.df <- osm_to_df(key = "leisure", value = "fitness_centre", type = "point")
playground.gg <- osm_point_plot(ff.map, play.df, "Fitness Centers")  
playground.gg  
play.df$object_id %>% unique() %>% length()
```

Fitness Center Observations: This is not reliable... not enough data points!

Use: No
```{r}

ff.map <- fairfax.gg()
play.df <- osm_to_df(key = "leisure", value = "sports_centre", type = "point")
playground.gg <- osm_point_plot(ff.map, play.df, "Sports Centers")  
playground.gg  
play.df$object_id %>% unique() %>% length()

```
Use: 

```{r}

ff.map <- fairfax.gg()
play.df <- osm_to_df(key = "leisure", value = "nature_reserve", type = "polygon")
playground.gg <- osm_poly_plot(ff.map, play.df, "Nature Reserves")  
playground.gg  
play.df$object_id %>% unique() %>% length()

``` 


```{r}

ff.map <- fairfax.gg()
play.df <- osm_to_df(key = "boundary", value = "national_park", type = "polygon")
playground.gg <- osm_poly_plot(ff.map, play.df, "National Parks")  
playground.gg  
play.df$object_id %>% unique() %>% length()

``` 







```{r}

ff.map <- fairfax.gg()
play.df <- osm_to_df(key = "leisure", value = "dog_park", type = "point")
playground.gg <- osm_point_plot(ff.map, play.df, "Dog Parks")  
playground.gg  
play.df$object_id %>% unique() %>% length()

``` 



```{r}

ff.map <- fairfax.gg()
play.df <- osm_to_df(key = "leisure", value = "garden", type = "point")
playground.gg <- osm_point_plot(ff.map, play.df, "Gardens")  
playground.gg  
play.df$object_id %>% unique() %>% length()

``` 


```{r}

ff.map <- fairfax.gg()
play.df <- osm_to_df(key = "leisure", value = "track", type = "point")
playground.gg <- osm_point_plot(ff.map, play.df, "Tracks")  
playground.gg  
play.df$object_id %>% unique() %>% length()

``` 
Track Obs: 



```{r}

ff.map <- fairfax.gg()
play.df <- osm_to_df(key = "leisure", value = "fitness_station", type = "point")
playground.gg <- osm_point_plot(ff.map, play.df, "Fitness Stations")  
playground.gg  
play.df$object_id %>% unique() %>% length()

``` 

Fitess Station Observations: Represent outdoor public fitness equipment; not enough points!
Use: No

