---
title: "Economic Vulnerability"
author: "VH"
date: "6/20/2019"
output: 
  github_document: default
  html_document: default
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
#Setting root directory
knitr::opts_knit$set(echo = TRUE,
                     root.dir = rprojroot::find_rstudio_root_file())

#To run the file you'll need to install all the packages here, in this order

#Packages
#install.packages("dplyr") # new  version
#install.packages(c("snakecase", "janitor")) #tidy names
#install.packages("pillar") #Dependency for patchwork
#devtools::install_github("thomasp85/patchwork") #For multi-panel plots
#install.packages(c("osmdata", "ggmap"))
#install.packages(c("sp", "sf"))
library(tidyverse) 
library(janitor)   #Clean names
library(patchwork) #Multi panel plots
library(osmdata)   #OSM
library(ggmap)     #Tidy gg maps
library(purrr)     #Map functions
library(sp)        #Spatial forms
library(sf)        #Spatial forms

#Controlling figure output in markdown
knitr::opts_chunk$set(
#  fig.height =   
  fig.width = 6,
#  fig.asp = .5,
  out.width = "90%",
#  out.height = 
  cache = FALSE
)

#Set Theme for ggplot2
theme_set(theme_bw() + theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom"))

#Set Scientific notation output for knitr
options(scipen = 999999)
```

##1. Define the Functions  

Before you do anything, restart your R session (under Session at the top). Then run the setup chunk above to initilize the packages you need and other output definitions for the .RMD document (don't worry about these for now). Run this entire chunk to define the functions for use below. You can verify that this worked if they show up in your global envirnment (top right), under functions.  

```{r read_osm, warning = FALSE}
#Fairfax read OSM to output tibble (tidyr df)
osm_to_df <- function(key = "leisure", value = "playground", type = "point") {
#Read in playground data
read.df <- getbb("fairfax county", format_out = "polygon") %>%
  opq() %>%
  add_osm_feature(key = key, value = value) %>%
  osmdata_sf()
    
    if (type == "point" & !is.null(read.df$osm_points)) {
      
      lat.long.df <- do.call(rbind, st_geometry(read.df$osm_points)) %>% 
        as_tibble() %>% 
        setNames(c("longitude", "latitude")) %>%
        mutate(object_id = 1:nrow(.) %>% as.factor()) %>%
        dplyr::select(object_id, everything())
      
    } else if (type == "polygon" & !is.null(read.df$osm_polygons)) {
      
      poly.list <- do.call(rbind, st_geometry(read.df$osm_polygons)) %>%
        lapply(as.tibble)
      npolys    <- length(poly.list)
      
      lat.long.df <- tibble(
        object_id = 1:npolys %>% as.factor(),
        coord.df  = poly.list
        ) %>%
        unnest(coord.df) %>%
        rename(
        longitude = lon,
        latitude  = lat
        )
      
    } else if (type == "line" & !is.null(read.df$osm_lines)) {
      
      line.list <- do.call(list, st_geometry(read.df$osm_lines)) %>%
        lapply(., rbind) %>%
        lapply(., as.tibble)
      nlines    <- length(line.list)
      
      lat.long.df <- tibble(
        object_id = 1:nlines %>% as.factor(),
        coord.df  = line.list
        ) %>%
        unnest(coord.df) %>%
        rename(
        longitude = lon,
        latitude  = lat
        )
    } else {
      ifelse(is.null(read.df$osm_points), stop("key/value incorrect or object type is empty"), 
             ifelse(!(type %in% c("point", "polygon", "line")), stop("type is not point, polygon, or line"),
                    stop("object type is empty, try another")))
    }
  
  return(lat.long.df)
}

#Base Map Function
fairfax.gg <- function() {
fairfax.box <- getbb("fairfax county")
fairfax.boundary <- getbb("fairfax county", format_out = "polygon") %>%
  as.tibble() %>%
  rename(longitude = `V1`, latitude = `V2`)

#Grab the map info (many varieties)
fairfax.map <- get_map(location = fairfax.box, source="stamen", maptype="watercolor", crop = TRUE)

#ggmap and ggplot map and boundary
ff.map <- ggmap(fairfax.map) +
  geom_polygon(data = fairfax.boundary, aes(x = longitude, y = latitude), colour = "black", size = 1, alpha = 0.1) +
  labs(
    x = "Longitude",
    y = "Latitude"
  )
  return(ff.map)
}

#Point Visualization
osm_point_plot <- function(ff.map, obj, value) {
  if (!is.data.frame(obj)) stop("OSM object is not a data frame")
  if (!is.ggplot(ff.map))  stop("Baseline Fairfax map is not a ggplot")
  ff.map +
  geom_point(data = obj, aes(x = longitude, y = latitude),
             size = 0.1, colour = "red", alpha = 0.25) + 
  geom_density_2d(data = obj, aes(x = longitude, y = latitude),
                  colour = "purple", alpha = 0.5) +
  labs(title = sprintf("Fairfax County %s", value))
  }

#Polygon Plot
osm_poly_plot <- function(ff.map, obj, value) {
  if (!is.data.frame(obj)) stop("OSM object is not a data frame")
  if (!is.ggplot(ff.map))  stop("Baseline Fairfax map is not a ggplot")
  ff.map +
  geom_polygon(data = obj, aes(x = longitude, y = latitude, group = object_id), 
               colour = "red", fill = "maroon", alpha = 0.5, size = 0.2) +
  labs(title = sprintf("Fairfax County %s", value))
  }

#Line Plot
osm_line_plot <- function(ff.map, obj, value) {
  if (!is.data.frame(obj)) stop("OSM object is not a data frame")
  if (!is.ggplot(ff.map))  stop("Baseline Fairfax map is not a ggplot")
  ff.map +
  geom_path(data = obj, aes(x = longitude, y = latitude, group = object_id), 
               colour = "red", alpha = 0.5, size = 0.5) +
  labs(title = sprintf("Fairfax County %s", value))
  }

#File Path Generator (for OSM specifically)
osm_filepath_gen <- function(data.folder, file.type, key, value, type) {
  date <- Sys.Date() %>% as.character() %>% str_replace_all("-", "_")
  file.type <- "csv"
  data.folder <- "original/osm"
  file.path <- sprintf("./data/%s/%s_%s.%s", 
                     data.folder,
                     date, 
                     paste(key, value, type, sep = "_"),
                     file.type)
  return(file.path)
}
```

