---
title: "Obesogenic Dimensional Analysis"
author: "DSPG Business Innovation Team"
date: "7/24/2019"
output: 
  github_document: default
  html_document: default
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
#Setting root directory
knitr::opts_knit$set(echo = TRUE,
                     root.dir = rprojroot::find_rstudio_root_file())

#Load Pacman for multiple package loads
if (!require("pacman")) install.packages(pacman)

#Need most updated dplyr if haven't already done this
#install.packages("dplyr") #need updated dplyr for sf objects

#Load all the good stuff
pacman::p_load(tidyverse, purrr, sf, sp, mapview, ggmap,
               patchwork, osmdata, mapview, traveltime,
               iterators, doParallel, foreach, parallel,
               geosphere, rgeos, raster, RapidPolygonLookup,
               corrplot, psych, factoextra)

#Controlling figure output in markdown
knitr::opts_chunk$set(
#  fig.height =   
  fig.width = 6,
#  fig.asp = .5,
  out.width = "90%",
#  out.height = 
  cache = TRUE
)

#Set Theme for ggplot2
theme_set(theme_bw() + theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom"))

#Set Scientific notation output for knitr
options(scipen = 999999)
```

##1. Read & Join the Obesogenic Data


Here we read the Fairfax Housing/OSM Binary Isochrone data, aggregate and nest by geography; read, clean, and nest by geography type the American Community Survey data (final Economic Vulnerability); and finally join by geography and write out a final Obesegenic Environment data frame for dimensional/factor analysis (and index construction).  

```{r message = FALSE, warning = FALSE}
#Read clean census track id's by manipulating the census tract number for matching join
census.tract.df <- read_csv("./data/working/ACS_final_index_2/index.csv") %>%
                     rename(geoid = Id2, census_tract = Geography) %>%
                     mutate(
                       geoid = census_tract %>% 
                         parse_number() %>% 
                         as.character() %>% 
                         ifelse(str_length(.) == 4, str_c(., ".00"), .) %>%
                         str_split("") %>%
                         map_chr(.x = ., ~.x[-5] %>% str_c(., collapse = ""))
                     )
                     
#Read in the Fairfax/OSM data
ffx.osm.df <- read_rds("./data/working/Fairfax_Housing_OSM_Joined/fairfax_osm_final.RDS") %>%
               mutate(
                 geoid = geoid %>% as.character() %>% str_sub(., 6, nchar(.)-1)
               )

#Bind together (3 observations do not have census tract identified & 3 census tracts not represented)
ffx.osm.df <- bind_rows(left_join(ffx.osm.df %>% filter(!is.na(geoid)), census.tract.df, by = "geoid"),
            ffx.osm.df %>% filter(is.na(geoid))) %>%
            dplyr::select(-parcel_id) %>%
            rename(
              highschool_district = highschool,
              supervisor_district = district
            ) %>%
            gather(key = id_type, 
                   value = geography,
                   c(census_tract, highschool_district, supervisor_district)) %>%
            mutate(id_type = as.factor(id_type)) %>%
            nest(-c(id_type, geography))

#Read in the Economic Vulnerability ACS/Housing Stock data
acs.df     <- read_csv("./data/working/ACS_final_index_2/07_22_2019_joined_acs_final.csv") %>%
  janitor::clean_names() %>%
  mutate(
    id_type   = as.factor(id_type) 
  ) %>%
  dplyr::select(id_type, geography, everything()) %>%
  nest(-c(id_type, geography)) 

#Join together for final Obesogenic Data: ACS, Fairfax/OSM (aggregated by geography, median or mean respectively)
obesity.df <- inner_join(ffx.osm.df, acs.df, by = c("id_type", "geography")) %>%
              rename(ffx_osm_df = data.x, acs_df = data.y) %>%
              unnest(acs_df) %>%
              mutate(
                ffx_osm_df = map(ffx_osm_df,
                                 ~.x %>%
                                   mutate(
                                     no_supermarket = !supermarket,
                                     no_playground  = !playground,
                                     no_sports_center = !sports_center,
                                     no_swimming_pool = !swimming_pool,
                                     no_team_sport    = !team_sport,
                                     no_track         = !track
                                   ) %>%
                                   dplyr::select(-c(supermarket, playground, sports_center,
                                             swimming_pool, team_sport, track))),
                ffx_osm_df = map(ffx_osm_df,
                                 ~.x %>%
                                   dplyr::select(alcohol:no_track) %>%
                                   map_df(mean))
              ) %>%
              unnest()

#Check counts
obesity.df %>% 
  group_by(id_type) %>%
  summarise(
    count = n()
  ) %>% knitr::kable()

#View Head and compare to make sure negativity mutation worked properly
#head(obesity.df) %>% knitr::kable()
#read_csv("./data/working/Obesogenic_final_data/2019_7_24_obesogenic_final.csv") %>% head()
```

```{r eval = FALSE, echo = FALSE}
#Write out a final .csv and nested .RDS object
write_csv(obesity.df, "./data/working/Obesogenic_final_data/2019_7_25_obesogenic_final.csv")
obesity.df %>% 
  nest(-id_type) %>%
  write_rds(., "./data/working/Obesogenic_final_data/2019_7_25_obesogenic_final_nested.RDS")
```

Here we were very meticulous about making sure everything is matched, aggregated, and assigned to the right geographic unit. After joining three seperate data sets, Fairfax Housing, OSM, and ACS; we have our final working Obesogenic Environment data (stored in the `"./data/working/Obesogenic_final_data/"` folder). Note that three observations are missing census tracts but are not missing supervisor or highschool district, while three census tracts are not present at all, thus there are three fewer total observations than in the original ACS units of geography (due to the three missing census tracts, as verified in the table above).  




##2. Dimensional Analysis  


####a. Singular Value Decomposition  


```{r}
#Read the data
obesity.df <- read_rds("./data/working/Obesogenic_final_data/2019_7_25_obesogenic_final_nested.RDS") %>%
              mutate(
                data = map(data,
                           ~.x %>%
                            dplyr::select(-c(minority, unmarried, single_parent, limited_english,
                               low_income, not_enrolled, no_vehicle, long_commute)))
              )

              
#Scale the data
obesity.scale.df <- obesity.df %>%
              mutate(
                data = data %>% map(.x = ., ~na.omit(.x) %>%
                                          dplyr::select(-geography) %>%
                                          mutate_if(is.numeric, function(x) {
                                            (x - mean(x))/sd(x)})),
                cor_data = data %>% map(.x = ., ~.x %>% cor(., method = "pearson")),
                pca_data = cor_data %>% map(princomp)
              ) 

#Singular Value Decomposition for Dimensions

#Corplots
obesity.scale.df$cor_data %>%
  map(.x = ., ~.x %>% corrplot(type = "upper", tl.col = "black", tl.cex = 0.75))


#Parallel analysis
obesity.scale.df$cor_data %>%
  map(.x = ., ~fa.parallel(.x))

#Scree Plot
obesity.scale.df$pca_data %>%
  map(.x = ., ~fviz_eig(.x))

#Feature Map
obesity.scale.df$pca_data %>%
  map(.x = ., ~.x %>% fviz_pca_var(.,
             col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE))
```



All evidence points to 3 distinct dimensions, however there is some discrepancy between 2 and 3 by supervisor district (the third plots); however for consistency we should probably fit a three factor model for all geographies. There is actually some interesting stuff going on at the census_tract level, if you look at the feature map in PCA space, there almost looks like there are actually 4 distinct groupings (but the groupings in the positive direction of Dimension 1 are less clear).   


