---
title: "OSM Obesity Data EDA"
author: "Quinton Neville"
date: "6/11/2019"
output:
  github_document: default
  html_document: default
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
#Setting root directory
knitr::opts_knit$set(echo = TRUE,
                     root.dir = rprojroot::find_rstudio_root_file())

#To run the file you'll need to install all the packages here, in this order

#Packages
#install.packages("dplyr") # new  version
#install.packages(c("snakecase", "janitor")) #tidy names
#devtools::install_github("thomasp85/patchwork") #For multi-panel plots
#install.packages("pillar") #Dependency for patchwork
#install.packages(c("osmdata", "ggmap", "sp"))
#install.packages(c("sp", "sf"))
library(tidyverse)
library(janitor)
library(patchwork)
library(ggridges) #Multi panel density plots
library(osmdata)
library(ggmap)
library(purrr)
library(sp)
library(sf)

#Controlling figure output in markdown
knitr::opts_chunk$set(
#  fig.height =   
  fig.width = 6,
#  fig.asp = .5,
  out.width = "90%",
#  out.height = 
  cache = TRUE
)

#Set Theme for ggplot2
theme_set(theme_bw() + theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom"))

#Set Scientific notation output for knitr
options(scipen = 999999)
```

#1. Explore OSM Data for Obesity Features

####a. Playgrounds  

Grab the data. Here, `get bb()` fetches the boundaries or box, `opq()` is the function which calls the  query through OSM's API, `add_osm_feature()` extracts the specific object(s) you are looking for and `osmdata_sf()` casts the object as a "simple feature object" (you can also use `osmdata_sp()` for quick plotting, see `CRAN`).   

Then, we access the longitude and lattitude coordinates in the `geometry` column of the `osm_points` object (1 per observation), and bind them together by row, cast as a tibble, and give it appropriate column names.  

```{r warning = FALSE}
#Read in playground data
play.df <- getbb("fairfax county", format_out = "polygon") %>%
  opq() %>%
  add_osm_feature(key = "leisure", value = "playground") %>%
  osmdata_sf()

#Pull latitude and longitude
play.lat.long.df <- do.call(rbind, st_geometry(play.df$osm_points)) %>% 
    as_tibble() %>% setNames(c("longitude", "latitude"))
```

Plot with `ggmap()` and `ggplot()`. First, get the boxed and ploygon area shapes of fairfax county. Use `get_map()` to pull the world map of your choice (there are tons of sources, and types to be standardized later). Then generate a `ggmap` of this object, draw the outline of fairfax county for reference to create the base Fairfax map object. Lastly, add on the features from the above `parks.df` to draw the desired points and density of parks in (an around) the county.  


```{r warning = FALSE, message = FALSE}
#Set Base Fairfax Map
#Boundary and outline of Fairfax County
fairfax.box <- getbb("fairfax county")
fairfax.boundary <- getbb("fairfax county", format_out = "polygon") %>%
  as.tibble() %>%
  rename(longitude = `V1`, latitude = `V2`)

#Grab the map info (many varieties)
fairfax.map <- get_map(location = fairfax.box, source="stamen", maptype="watercolor", crop = TRUE)

#ggmap and ggplot map and boundary
ff.map <- ggmap(fairfax.map) +
  geom_polygon(data = fairfax.boundary, aes(x = longitude, y = latitude), colour = "black", size = 1, alpha = 0.1) +
  labs(
    x = "Longitude",
    y = "Latitude"
  )

#Add locations of playgrounds
playground.gg <- ff.map +
  geom_point(data = play.lat.long.df, aes(x = longitude, y = latitude),
             size = 0.1, colour = "red", alpha = 0.25) + 
  geom_density_2d(data = play.lat.long.df, aes(x = longitude, y = latitude),
                  colour = "purple", alpha = 0.5) +
  labs(title = "Fairfax County Playgrounds")

#Visualize
playground.gg
```




####b. Parks  

Repeat the process in a., except now we are interested in polygons or park areas. To do so we apply a slightly different approach but basically get the polygon outline (tibble of lat/long) for each park (1703) in a nested (a tibble of tibbles) frame, then `unnest()` it to obtain the ultimate long frame used for plotting each park area.  


```{r park, warning = FALSE}
#Read in playground data
park.df <- getbb("fairfax county", format_out = "polygon") %>%
  opq() %>%
  add_osm_feature(key = "leisure", value = "park") %>%
  osmdata_sf()

#Obtain list of tibbles for each park's lat, long
park.list <- do.call(rbind, st_geometry(park.df$osm_polygons)) %>%
  lapply(as.tibble)

#Grab length (number) of parks
nparks <- length(park.list)

#Comine w/factor park identifier and unnest (make long for plotting)
park.lat.long.df <- tibble(
  park     = 1:nparks %>% as.factor(),
  coord.df = park.list
) %>%
  unnest(coord.df) %>%
  rename(
    longitude = lon,
    latitude  = lat
  )
```

Plot with `ggmap()`, same idea as a. above. 
```{r park_plot, warning = FALSE}
#Add polygons of parks
park.gg <- ff.map +
  geom_polygon(data = park.lat.long.df, aes(x = longitude, y = latitude, group = park), colour = "red", fill = "maroon", alpha = 0.5, size = 0.2) +
  labs(title = "Fairfax County Park Areas")

park.gg
```

####c. Patchwork to Panel display

This package `patchwork` has really intuitive ways to panel `gg objects` or `globs`, look it up for further instructions but `+` puts things side by side, `/` one over the other, etc.  

```{r patchwork, warning = FALSE}
playground.gg + park.gg
```